We would like to express our sincere gratitude to our supervisor, Dr. Péter Kovács, for his invaluable guidance and support throughout this research project. We are deeply thankful for the opportunity to work on compression-aware ECG denoising and to be part of the ELTE Modelling Lab. His trust in allowing us to pursue this work as a TDK thesis project has been instrumental in our academic development.

Since the beginning of this project, Dr. Kovács's expert guidance, constructive feedback, and profound knowledge in signal processing and deep learning have been essential in shaping both the technical direction and scientific rigor of this work. We are particularly grateful for his regular supervision meetings, his patience during the iterative development process, and his continuous encouragement, especially during the more demanding phases of experimentation and analysis. His commitment to methodological clarity and scientific excellence has significantly influenced our approach to research and problem-solving.

We would also like to extend our heartfelt appreciation to our families and friends for their unwavering support, encouragement, and understanding throughout this research journey. Their patience and belief in our work have been a constant source of motivation.

Finally, we gratefully acknowledge the Stipendium Hungaricum Scholarship Programme for providing the financial support that made our studies in Hungary possible. Without this opportunity, the pursuit of this research would not have been feasible.

